### 2025-12-13 disser/part2.tex
disser

Для объяснения наблюдаемых профилей рентгеновской яркости в близи первой реколлимационной ударной волны (областях `Head' и `Cone') мы использовали следующую многокомпонентную модель магнитного поля - сумма описанных выше двух турбулентных полей, с равными на фронте значениями магнитного поля $7.5$ мкГс и характерными масштабами $3\times10^{17}$ см для короткомасштабной неустойчивости и $3\times10^{19}$ см для крупномасштабной. Так же считалось, что среднеквадратичное перестает спадать с удалением от фронта при достижении значения $8$ мкГс. Скорость течения за фронтом ударной волны бралась равной $v = 10.3\times10^8$ см/с (согласованно с Монте-Карло моделированием спектра ускоренных частиц). Результаты моделирования синхротронного излучения, а так же наблюдаетельные данные \cite{Safi-Harb2022} представлены на рисунке \ref{synchrotron}. Полученные спектры согласуются с данными наблюдений

@@

\subsection{Гамма излучение}
In this Section, we model the gamma radiation of W50, which is observed by H.E.S.S. \cite{HESS2024SS433} and LHAASO \cite{LHAASO2024SS433}. The leptonic model assumes that it is produced via inverse Compton scattering of electrons accelerated by recollimation shocks in the collimated outflow.

In Fig. \ref{compton}, the red curve shows the gamma radiation produced within the outflow by accelerated VHE electrons as they are advected through the downstream of the first shock to the second one. This corresponds to pure advection of VHE electrons along the X-ray outflow with the turbulent magnetic field and the distribution function of electrons, which were used to simulate the synchrotron X-ray radiation discussed above. The red curve is significantly below the observational data, and it has a different spectral shape at low energies. This is because the radiating electrons in the X-ray jet are in the thin target regime.
On the other hand, the electrons accelerated at the first shock diffuse in the direction transverse to the X-ray outflow. These electrons left the axial X-ray outflow and diffused through the surrounding cocoon, comprising a spatially broad component. If their diffusion coefficient in the cocoon is $\sim 10^{28} см^2/c$ at 100 TeV, they would radiate gamma-rays in the thick target regime in a wide cocoon of a size $\sim$ 30 pc around the axial jet. Electrons that have flown through the X-ray axial jet with a magnetic field amplified by cosmic ray-driven instabilities at the shock and escape into the cocoon downstream would also radiate in the thick regime. In Fig. \ref{compton}, the gamma-ray spectrum integrated over the wide cocoon is shown by the blue curve, which is in reasonably good agreement with the LHAASO data points taken from \cite{LHAASO2024SS433}. The gamma-ray spectrum integrated over a somewhat narrower region of the cocoon corresponds to the intermediate regime between the thin and the thick targets. It is shown by a brown curve in Fig.~\ref{compton} and is compared with the data reported by H.E.S.S. \cite{HESS2024SS433}.

The LHAASO images above 100 TeV \cite{LHAASO2024SS433} show a more extended emission region compared to the lower energy bands. This morphology can be understood within our diffusive shock acceleration model, where the highest energy particles escape into the shock upstream, where the low magnetic field allows the escaped VHE leptons to form an extended gamma-ray emission region by the inverse Compton process.

In our model, $\sim$ 30 $\%$ of the electrons accelerated at the shock front had escaped from the jet and diffused in the cocoon. This allows us to explain the observational data from LHAASO as shown in Figure \ref{compton}. H.E.S.S. data fall between the model for LHAASO and radiation from the jet itself, which can be explained by a smaller region size, observed by H.E.S.S.

There are recent studies of the acceleration of PeV protons in microquasars \citep{2024A&A...691A..93A,2024arXiv241108762P,2025arXiv250510620K}.  A spherically symmetric semi-analytical model of particle acceleration to energies of several PeV by the termination wind shock was discussed in \cite{2024arXiv241108762P}. The acceleration of particles in the shear flows of jet-cocoon-structured microquasars is suggested in \citep{2025arXiv250620193Z}. Shear acceleration in the model requires the injection of particles pre-accelerated up to TeV regime energies. Hadronic interactions of accelerated protons are producing a substantial fraction of the observed VHE gamma-ray emission in the models \citep{2024arXiv241108762P,2025arXiv250510620K}.

The recollimation shocks in the diffusive shock acceleration model considered here efficiently accelerate protons up to PeV energies, as is seen in Fig. \ref{pdf_shock_Q_esc}.
In our model, $\gsim$ 10\% of the shocked jet power is transferred to accelerated protons while the VHE emission comes from the inverse Compton radiation of VHE electrons accelerated at shocks, producing polarized synchrotron X-ray emission of the extended X-ray jet. The leptonic gamma-ray emission is produced in the thick target regime within the cocoon surrounding the extended X-ray jet. 
The VHE gamma rays can be produced by hadronic interactions of the protons with ambient matter. The very low number density of the plasma in the extended jet and in the surrounding cocoon would require a very small diffusion coefficient for the accelerated VHE protons to contribute to the detected gamma-ray emission from the cocoon, given by the gamma-ray significance maps of H.E.S.S. and LHAASO. Therefore, most of the accelerated protons would escape the accelerator in the SS433/W50 extended X-ray jet and then contribute to the gamma-ray emission from the denser gas in the vicinity of W50 and supply the galactic cosmic-ray population.   
\section{Выводы} \label{SS433conclision}

>>>>

Для объяснения наблюдаемых профилей рентгеновской яркости в близи первой реколлимационной ударной волны (областях `Head' и `Cone') мы использовали следующую многокомпонентную модель магнитного поля - сумма описанных выше двух турбулентных полей, с равными на фронте значениями магнитного поля $7.5$ мкГс и характерными масштабами $3\times10^{17}$ см для короткомасштабной неустойчивости и $3\times10^{19}$ см для крупномасштабной. Так же считалось, что среднеквадратичное перестает спадать с удалением от фронта при достижении значения $8$ мкГс. Скорость течения за фронтом ударной волны бралась равной $v = 10.3\times10^8$ см/с (согласованно с Монте-Карло моделированием спектра ускоренных частиц). Результаты моделирования синхротронного излучения, а так же наблюдаетельные данные \cite{Safi-Harb2022} представлены на рисунке \ref{synchrotron}. Полученные спектры согласуются с данными наблюдений. Уровень рентгеновского излучение так же позволяет оценить отношение инжекций электронов и протонов в процесс ускорения. Как уже говорилось выше, Монте-Карло моделирование не способно корректно расчитывать инжекцию электронов, и нормировка нетеплового хвоста их распределения произвольна. Вычисление абсолютной величины потоков синхротронного излучение позволяет определить абсолютную величину функции распределения электронов. В данном случае правильная величина потоков получается при отношении инжекций $3.3\times10^{-3}$, что соответствует теоретическим оценкам из работы \cite{ParkElectronIonShock}.

@@

\subsection{Гамма излучение}

Туманность W50 является ярким источником гамма-излучения. Она наблюдалась телескопами HAWC \cite{Abeysekara2018}, H.E.S.S \cite{HESS2024SS433} и LHAASO \cite{LHAASO2024SS433}. Мы предположили лептонную модель происхождения данного излучения - в результате обратного комптоновского рассеяния электронов, ускоренных на реколлимационных ударных волнах.

Спектры, расчитанные в различных моделях, показаны на рисунке \ref{compton}. Красная кривая соответствует излучению ускоренных электронов, конвективно переносимых вдоль джета от первой реколлимационной ударной волны, как описано в предыдущем разделе. С учетом нормировки, полученной из оценки рентгеновского излучения, оказывается, что излучения этих электронов не хватает для объяснения наблюдаемого гамма излучения. Так же видно, что на низких энергиях степенной индекс намного жестче, чем наблюдаемое значение, что указывает на то, что наблюдаемое излучение происходит в режиме толстой мишени даже на низких энергиях. Отсюда можно предположить, что часть ускоренных электронов покидает джет за счет поперечной диффузии и попадает в область слабого магнитного поля в коконе, где излучает преимущественно за счет обратного комптоновского рассеяния. Если коэффициент диффузии на энергиях около 100 ТэВ меньше $10^{28} см^2/c$, то такие электроны могут удерживаться в коконе шириной около 30 пк на временах порядка десятков тысяч лет и излучать в режиме толстой мишени, теряя всю свою энергию на рождение гамма квантов. Если доля таких электронов, перешедших из джета в кокон, не теряя энергию на синхротронное излучение составит $30\%$ от общего числа ускоренных электронов, то суммарное излучение, показанное розовой кривой на рисунке \ref{compton} позволит объяснить данные H.E.S.S и высокоэнергичную часть данных LHAASO.

\begin{figure}
	\includegraphics[width=0.8\textwidth]{W50compton3.png}
	\caption{Моделированный спектр гамма излучения W50 и наблюдательные данные\cite{HESS2024SS433, LHAASO2024SS433}. }
	\label{compton}
\end{figure}

Для объяснения подъема на низкоэнергичной части спектра LHAASO предположим, что электроны, полностью прошедшие вдоль джета и потерявшие значительную часть энергии на синхротронное излучение, в конце концов тоже попадают в кокон и свободно диффундируют в нем, так же излучая обратным комптоновским рассеянием в режиме толстой мишени. Сумарный спектр трех компонент даст голубую кривую на рисунке \ref{compton}, хорошо совпадающую с данными LHAASO. Предположении о наличии большего числа электронов, видимых LHAASO чем HESS, косвенно подтверждается большими размерами областей со значимым сигналом, которые видит LHAASO.

\section{Выводы} \label{SS433conclision}
